---
title: "Repoducible Research Project 1"
author: "Kaihua Hou"
date: "7/2/2020"
output: html_document
---

``` {r setOptions, echo = FALSE}
library(knitr)
library(GlobalOptions)
opts_chunk$set(cache = TRUE)
```
      
### 1. Loading and preprocessing the data
#### reading the csv file into R

``` {r readingData}
activityZip <- download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", destfile="./activity.zip")
activity <- unzip("activity.zip")
activityData <- read.csv("activity.csv", header = TRUE)
```

#### processing the values in the data
```{r naRemoval, results = "hide"}
activityData <- na.omit(activityData, cols="steps")
as.numeric(activityData$steps)
```


### 2. Mean total number of steps taken per day
#### making a histogram of the total number of steps taken each day

``` {r histogram}
totalSteps <- aggregate(steps ~ date, activityData, FUN=sum)
hist(totalSteps$steps,
     main = "Total Steps per Day",
     xlab = "Number of Steps")
```

#### Calculate and report the mean and median total number of steps taken per day
```{r meanAndMedian}
mean(totalSteps$steps, na.rm = TRUE)
median(totalSteps$steps, na.rm = TRUE)
```

### 3. The average daily activity pattern
## steps taken, averaged acoss all days
``` {r }
library(ggplot2)
meanStepsByInt <- aggregate(steps ~ interval, activityData, mean)
ggplot(data = meanStepsByInt, aes(x = interval, y = steps)) +
  geom_line() +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps")
```

### maximum number of steps among the 5-minute intervals
``` {r }
maxInt <- meanStepsByInt[which.max(meanStepsByInt$steps),]
head(totalSteps)
```


### 4. Imputing Missing Values 
#### report missing values
``` {r }
missingVals <- is.na(activityData$steps)
```

#### creating a new dataset that is equal to the original dataset but with the missing data filled in, and making a histogram of the total number of steps taken each day
``` {r}
imp_activityData <- transform(activityData,
                              steps = ifelse(is.na(activityData$steps),
                                             meanStepsByInt$steps[match(activityData$interval, meanStepsByInt$interval)], activityData$steps))

impStepsByInt <- aggregate(steps ~ date, imp_activityData, FUN=sum)
hist(impStepsByInt$steps,
     main = "Imputed Number of Steps Per Day",
     xlab = "Number of Steps")
```


### Differences in activity patterns between weekdays and weekends
#### Make a panel plot containnig a time-series plot of the 5-minute interval and the average number of steps taken across all weekdays or weekends
```{r}
DayType <- function(date) {
  day <- weekdays(date)
  if (day %in% c('Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday'))
      return ("weekeday")
  else if (day %in% c('Saturday', 'Sunday'))
      return ("weekend")
  else
      stop ("Invalid Date Format.")
}
imp_activityData$date <- as.Date(imp_activityData$date)
imp_activityData$day <- sapply(imp_activityData$date, FUN = DayType)

meanStepsByDay <- aggregate(steps ~ interval + day, imp_activityData, mean)
ggplot(data = meanStepsByDay, aes(x = interval, y = steps)) + 
  geom_line() +
  facet_grid(day ~ .) +
  ggtitle("Average Daily Activity Pattern") +
  xlab("5-minute Interval") +
  ylab("Average Number of Steps")
```


